Data Exploration quad maps: IUCN and AquaMaps

### 2016-03-31 data_explore/data_explore.Rmd
added code to ditch IUCN risk colors

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc, 
 #                               size = lg_area_pct,
                               color = popn_category)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#d01c8b") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey30')) + 
#   scale_size_continuous(limits = c(0, 100),
#                         breaks = c( 0,    25,    50,    75,    100), 
#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
#       title = 'Distribution alignment vs Extent alignment',
#       size = '% ocean range',
       color = 'Extinction\nRisk')

@@

ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
       width = 17.8, height = 9, units = 'cm')

@@

>>>>

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc)) + #, 
 #                               size = lg_area_pct,
 #                              color = popn_category)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#d01c8b") + 
  geom_point(color = 'grey40', alpha = .6)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
#   scale_colour_manual(values = c('LC' = 'green4', 
#                                  'NT' = 'greenyellow', 
#                                  'VU' = 'yellow', 
#                                  'EN' = 'orange', 
#                                  'CR' = 'orangered2',
#                                  'DD' = 'grey30')) + 
#   scale_size_continuous(limits = c(0, 100),
#                         breaks = c( 0,    25,    50,    75,    100), 
#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)')) #, 
#       title = 'Distribution alignment vs Extent alignment',
#       size = '% ocean range',
 #      color = 'Extinction\nRisk')

@@

ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
       width = 17.8, height = 9, units = 'cm')

@@

``` {r risk quads linear model}

spp_risk_df <- spp_list_quads %>%
  select(iucn_sid, spp_group = spp_group_text, popn_category, category_score, quad) %>%
  mutate(quad_good = ifelse(quad == 'q1', 1, 0))

risk_mdl1 <- glm(as.factor(quad_good) ~ category_score + spp_group, 
                 family = binomial,
                 data = spp_risk_df)

risk_mdl2 <- glm(as.factor(quad_good) ~ category_score, 
                 family = binomial,
                 data = spp_risk_df)

risk_mdl3 <- glm(as.factor(quad_good) ~ spp_group, 
                 family = binomial,
                 data = spp_risk_df)

BIC(risk_mdl1, risk_mdl2, risk_mdl3)
summary(risk_mdl1)
summary(risk_mdl2)
summary(risk_mdl1)

binomial logistic regression model

# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q1'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q2'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q3'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q4'])

### 2016-03-31 data_explore/data_explore_quadmaps.Rmd
added code to ditch IUCN risk colors

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### this file created in data_prep
spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)

@@

@@

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc, 
                               color = popn_category)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#d01c8b") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey30')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

### here are individual species, identified and namean on the plot:
scatter_quadplot <- scatter_quadplot +
  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
  geom_text(data = spp_list_label, 
            aes(label = sciname_label), 
            color = 'grey30',
            fontface = 'bold.italic',
            size = 3,
            hjust = .75,
            vjust = 1.5)

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
       color = 'Extinction\nRisk')

>>>>

@@

spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)

spp_list <- spp_list %>% 
  mutate(popn_category = ifelse(popn_category == 'LC', 'Least concern', popn_category),
         popn_category = ifelse(popn_category == 'NT', 'Near threatened', popn_category),
         popn_category = ifelse(popn_category == 'VU', 'Vulnerable', popn_category),
         popn_category = ifelse(popn_category == 'EN', 'Endangered', popn_category),
         popn_category = ifelse(popn_category == 'CR', 'Critically endangered', popn_category),
         popn_category = ifelse(popn_category == 'DD', 'Data deficient', popn_category)) %>%
  transform(popn_category = reorder(popn_category, category_score))

@@

### mongo plot time
scatter_quadplot <- ggplot(spp_list,
                           aes(x = area_ratio, 
                               y = sm_perc, 
                               color = popn_category)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#d01c8b") + 
  geom_point(alpha = .5)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
  scale_colour_manual(values = c('Least concern' = 'green4', 
                                 'Near threatened' = 'greenyellow', 
                                 'Vulnerable' = 'yellow', 
                                 'Endangered' = 'orange', 
                                 'Critically endangered' = 'orangered2',
                                 'Data deficient' = 'grey30')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

### here are individual species, identified and namean on the plot:
scatter_quadplot <- scatter_quadplot +
  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
  geom_text(data = spp_list_label, 
            aes(label = sciname_label), 
            color = 'grey30',
            fontface = 'bold.italic',
            size = 2.5,
            hjust = .75,
            vjust = 1.5)

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
       color = 'Extinction Risk')
